An Adolescent Mini-Neptune in the Kepler Field

### 2021-08-13 paper/ms.tex
update ms

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby ($d\approx
  333\,{\rm pc}$) and it resembles the Sun.  Combining kinematic
  information from Gaia, stellar rotation periods from TESS, and Li
  abundances from ground-based spectra, we show that Kepler\,1627 is a
  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.  To our
  knowledge, Kepler\,1627Ab is the youngest planet with a precise age
  yet found by the main Kepler mission.  From both unresolved
  starspot-crossing events in the Kepler photometry and from a
  ground-based Keck/HIRES transit we find that the planet is likely on
  an orbit nearly aligned with the stellar spin axis.  We also find
  from the short-cadence Kepler photometry that although the flare
  arrival time distribution is probably Poissonian, there may be an
  excess of flares separated by the orbital and synodic periods, and
  linear combinations thereof.  The $\delta$~Lyr cluster is one of
  many stellar groups whose properties have been clarified using the
  Gaia data; many of the other new clusters also overlap with
  confirmed and candidate exoplanets.  Further studies that combine
  Gaia kinematics and TESS rotation periods are poised to expand
  the census of age-dated planets, and our understanding of planetary
  evolution.
\end{abstract}

@@

In brief , the age of the \cn\ cluster based on the empirical
fitting procedure described in Section~\ref{sec:cmd} is
$35^{+6}_{-5}$\,Myr.  This procedure hinges on the age adopted for
IC\,2602, since it is the closest reference cluster in the
color-magnitude diagram.  Gyrochronologically, while most of the G and
K dwarfs in the \cn\ have not yet converged to a slow sequence, their
rotation distribution and that of the F dwarfs is consistent with the
same age.  The rotation periods of other 30\,Myr to 50\,Myr clusters
({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
\citep{douglas_stephanie_t_2021_5131306}.  The most precise age for \sn\ and its
planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's
spatial, kinematic, and rotational connection to this larger ensemble,
and the assumption that the planet formed very quickly after the star.

%TODO:  check Teff here once you have it.
Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet.  The utility
of this element for age dating young stars has been reviewed by
\citet{soderblom_ages_2014}.  We acquired an iodine-free spectrum from
Keck/HIRES on the night of 26 March 2021 using the standard setup and
reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we measure ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}\,$m\AA.   Note that this value does not
correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given $T_{\rm
eff} \approx 5500\,{\rm K}$, this measurement is also in agreement
with an age for the star between 30\,Myr and 50\,Myr (see {\it e.g.},
Figure~9 of \citealt{bouma_2021_ngc2516}, or the data from
\citealt{randich_gaiaeso_2018}).

@@

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f16.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
    {\it Top:} 
    The time of each Kepler transit was measured, along with the local slope
    of the light curve.  The two quantities appear anti-correlated,
    which is most easily explained by starspot crossings during the
    first (second) half of transit inducing a positive (negative) TTV,
    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
    units along the abscissa are most easily understood by considering
    that the flux from the star changes by $\sim$60\,ppt per half
    rotation period ($\sim$1.3\,days).
        {\it Bottom:} A Keck/HIRES transit was observed on 7 Aug 2021 UT,
    simultaneous with MUSCAT3.  A quadratic baseline trend was allowed
    in the fit.
    The resulting best-fit sky-projected obliquity is listed in red.
    \label{fig:obliquity}
	}
\end{figure}

@@

{\bf todo: Kento to check numbers on per-transit analysis} 
The second line of evidence comes from our transit timing analysis. We
isolated each of the 144 observed transits to within $\pm4.5$\,hr of
each transit, and fitted each window with both {\it i)} a local
second-order polynomial and transit, and {\it ii)} a local linear
trend.  We let the mid-time of each transit float, and then calculated
the residual between the measured mid-time and that of a perfectly
periodic orbit.  This residual, the ``transit timing variation''
(TTV), is plotted in Figure~\ref{fig:obliquity} against the local
linear slope.  Only a handful of Kepler Objects of Interest have shown
this correlation \citep{holczer_time_2015}, which is most readily
interpreted as a TTV induced by unresolved starspot crossings
\citep{mazeh_time_2015}.  This is only possible if the
planet transits the primary star, which excludes the
aforementioned false positive scenarios.  It also implies that the
planet's orbit is prograde.  The latter interpretation assumes that
the dominant photometric variability is induced by dark spots, rather
than bright faculae.  Given the observed transition of Sun-like
stellar variability from spot to faculae-dominated regimes, we expect
this inference to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

The final line of evidence confirming \pn\ comes from a spectroscopic
transit we measured from Keck/HIRES during the night of 7 Aug 2021
(UT) using the iodine cell for the wavelength calibration.  We
scheduled the transit using the ephemeris of
\citet{holczer_transit_2016}, which we verified using the available
TESS data.  We also acquired simultaneous multi-band photometry using
MUSCAT3, to help interpret the radial velocities \citep{Narita_2020}.
The lower panels in Figure~\ref{fig:obliquity} show the results.  The
photometry shows the expected starspot trend, along with the transit
and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
The spectroscopic data show significant scatter, though a significant
increase in radial velocity is seen shortly after the expected
ingress.  We fitted the RVs using the \citet{hirano_analytic_2010} and
\citet{hirano_2011} models for the Rossiter-McLaughlin effect, and
considered the cases of a linear and quadratic trend.
% TODO finish this paragraph
We found $\chi^2_{\rm red} = X.XX$ for the best-fitting linear case, 
and
$\chi^2_{\rm red} = X.XX$ for the best-fitting quadratic, with a
$\Delta {\rm BIC} = Y.YY$ between the two, providing {\bf (very)
strong} evidence to prefer the latter.

@@

Ultimately, our age measurement for \pn\ was enabled by identifying
its connection to the \cn\ using the Gaia data.  The Supplementary
Table~\ref{tab:v05} enables similar crossmatches to already known
planets, as well as forthcoming planets from the TESS mission
\citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
approach would be to identify kinematic associations around exoplanet
host stars using the Gaia data, and to then verify these associations
with ancilliary rotation periods and spectroscopy
\citep{tofflemire_tess_2021}.

@@

\input{model_posterior_table.tex}

@@

L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
transit fitting, to R{.}~Kerr for kindly providing us with the
\citet{kerr_stars_2021} membership list prior to its publication.
% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
% that encouraged this analysis.
%
L.G.B{.} acknowledges support by the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
L.G.B{.} was also supported by a Charlotte Elizabeth Procter
Fellowship from Princeton University.
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%

%
This study was based in part on observations at Cerro Tololo
Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
Association of Universities for Research in Astronomy (AURA) under a
cooperative agreement with the National Science Foundation.
%
%
This research has made use of the Keck Observatory Archive (KOA),
which is operated by the W. M. Keck Observatory and the NASA Exoplanet
Science Institute (NExScI), under contract with the National
Aeronautics and Space Administration.
%

@@

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.6\textwidth]{TEMP_binarity_ao.png}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Keck NIRC2 AO image of Kepler 1627A and Kepler 1627B.}  
    {\bf SCALEBAR DENOTES...}
    {\bf North is up, East is left}.
    \label{fig:ao}
	}
\end{figure*}

>>>>

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Combining
  kinematic information from Gaia, stellar rotation periods from TESS,
  and Li abundances from ground-based spectra, we show that
  Kepler\,1627 is a member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr
  cluster.  To our knowledge, Kepler\,1627Ab is the youngest planet
  with a precise age yet found by the main Kepler mission.  From both
  unresolved starspot-crossing events in the Kepler photometry and
  from a ground-based Keck/HIRES transit we find that the planet is
  likely on an orbit nearly aligned with the stellar spin axis.  We
  also find from the short-cadence Kepler photometry that although the
  flare arrival time distribution is probably Poissonian, there may be
  an excess of flares separated by the orbital and synodic periods,
  and linear combinations thereof.  The $\delta$~Lyr cluster is one of
  many stellar groups whose properties have been clarified using the
  Gaia data; many of the other new clusters also overlap with
  confirmed and candidate exoplanets.  Further studies that combine
  Gaia kinematics and TESS rotation periods are poised to expand the
  census of age-dated planets, and our understanding of planetary
  evolution.
\end{abstract}

@@

In brief, the age of the \cn\ cluster based on the empirical
fitting procedure described in Section~\ref{sec:cmd} is
$35^{+6}_{-5}$\,Myr.  This procedure hinges on the age adopted for
IC\,2602, since it is the closest reference cluster in the
color-magnitude diagram.  Gyrochronologically, while most of the G and
K dwarfs in the \cn\ have not yet converged to a slow sequence, their
rotation distribution and that of the F dwarfs is consistent with the
same age.  The rotation periods of other 30\,Myr to 50\,Myr clusters
({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
\citep{douglas_stephanie_t_2021_5131306}.  The most precise age for \sn\ and its
planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's
spatial, kinematic, and rotational connection to this larger ensemble,
and the assumption that the planet formed very quickly after the star.

Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet.  The utility
of this element for age dating young stars has been reviewed by
\citet{soderblom_ages_2014}.  We acquired an iodine-free spectrum from
Keck/HIRES on the night of 26 March 2021 using the standard setup and
reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we measure ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}\,$m\AA.   Note that this value does not
correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given the stellar
parameters of \sn\ (Table~\ref{tab:starparams}), most notably its
effective temperature, this measurement is also in agreement with an
age for the star between 30\,Myr and 50\,Myr (see {\it e.g.}, Figure~9
of \citealt{bouma_2021_ngc2516}, or the data from
\citealt{randich_gaiaeso_2018}).

@@

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f16.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
    {\it Top:} 
    The time of each Kepler transit was measured, along with the local slope
    of the light curve.  The two quantities appear anti-correlated,
    which is most easily explained by starspot crossings during the
    first (second) half of transit inducing a positive (negative) TTV,
    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
    units along the abscissa are most easily understood by considering
    that the flux from the star changes by $\sim$60\,ppt per half
    rotation period ($\sim$1.3\,days).
        {\it Bottom:} Keck/HIRES transit of 7 Aug 2021 UT,
    observed simultaneously with MUSCAT3.  A quadratic baseline trend was allowed
    in the fit.
    The resulting best-fit sky-projected obliquity is listed in red.
    \label{fig:obliquity}
	}
\end{figure}

@@

% TODO: Kento Masuda, please verify
The second line of evidence comes from our transit timing analysis. We
isolated each of the 144 observed transits to within $\pm4.5$\,hr of
each transit, and fitted each window with both {\it i)} a local
second-order polynomial and transit, and {\it ii)} a local linear
trend.  We let the mid-time of each transit float, and then calculated
the residual between the measured mid-time and that of a perfectly
periodic orbit.  This residual, the ``transit timing variation''
(TTV), is plotted in Figure~\ref{fig:obliquity} against the local
linear slope.  A significant negative slope of $-0.0564 \pm 0.0098\
{\rm ppt\,day}^{-1}$ is observed.  Only a handful of Kepler Objects of
Interest have shown this correlation \citep{holczer_time_2015}, which
is most readily interpreted as a TTV induced by unresolved starspot
crossings \citep{mazeh_time_2015}.  This is only possible if the
planet transits the primary star, which excludes the aforementioned
false positive scenarios.  It also implies that the planet's orbit is
prograde.  The latter interpretation assumes that the dominant
photometric variability is induced by dark spots, rather than bright
faculae.  Given the observed transition of Sun-like stellar
variability from spot to faculae-dominated regimes, we expect this
inference to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

The final line of evidence confirming \pn\ comes from a spectroscopic
transit we measured from Keck/HIRES during the night of 7 Aug 2021
(UT), using the iodine cell for the wavelength calibration.  We
scheduled the transit using the ephemeris of
\citet{holczer_transit_2016}, which we verified using the available
TESS data.  We also acquired simultaneous multi-band photometry using
MUSCAT3, to help interpret the radial velocities \citep{Narita_2020}.
The lower panels in Figure~\ref{fig:obliquity} show the results.  The
photometry shows the expected starspot trend, along with the transit
and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
The spectroscopic data show significant scatter, though a significant
increase in radial velocity is seen shortly after the expected
ingress.  We fitted the RVs using the \citet{hirano_analytic_2010} and
\citet{hirano_2011} models for the Rossiter-McLaughlin effect, and
considered the cases of a linear and quadratic trend.
% TODO finish this paragraph
We found $\chi^2_{\rm red} = X.XX$ for the best-fitting linear case, 
and
$\chi^2_{\rm red} = X.XX$ for the best-fitting quadratic, with a
$\Delta {\rm BIC} = Y.YY$ between the two, providing {\bf (very)
strong} evidence to prefer the latter.

@@

Ultimately, our age measurement for \pn\ was enabled by identifying
its connection to the \cn\ using the Gaia data.  The Supplementary
Table~\ref{tab:v05} enables similar crossmatches to already known
planets, as well as forthcoming planets from the TESS mission
\citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
approach would be to identify kinematic associations around exoplanet
host stars using tangential velocities from Gaia, and to then verify these associations
with ancilliary rotation periods and spectroscopy
\citep{tofflemire_tess_2021}.

@@

@@

L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
transit fitting, to R{.}~Kerr for kindly providing us with the
\citet{kerr_stars_2021} membership list prior to its publication.
L.G.B{.}, L.H{.}, and N.N{.} are also grateful to K{.}~Collins for
helping resolve the scheduling conflict that would have otherwise
prevented the MUSCAT3 observations.
% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
% that encouraged this analysis.
%
L.G.B{.} acknowledges support by the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
L.G.B{.} was also supported by a Charlotte Elizabeth Procter
Fellowship from Princeton University.
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%

%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
This research has made use of the Keck Observatory Archive (KOA),
which is operated by the W. M. Keck Observatory and the NASA Exoplanet
Science Institute (NExScI), under contract with the National
Aeronautics and Space Administration.
We also thank
the Keck Observatory staff, particularly {\bf A.~Doe},
for support of HIRES and remote observing.  We recognize
the importance
that the summit of Maunakea has within the indigenous
Hawaiian community, and are deeply grateful to
have the opportunity to conduct observations from this
mountain.
%

@@

\input{starparams.tex}
\input{model_posterior_table.tex}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.6\textwidth]{Kepler1627_20210624_562_832_final.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Keck NIRC2 AO image of Kepler 1627A and Kepler 1627B.}  
    {\bf SCALEBAR DENOTES...}
    {\bf North is up, East is left}.
    \label{fig:ao}
	}
\end{figure*}
